---
title: "NY Housing"
output: word_document
---

**1.a** I plan on looking at how price is determined using a multivariate regression taking real estate data into account. The features I want to take into account are gross square feet, the neighborhood, and building class. I also want to look at the relationship of sale-date and price across the entire dataset to identify patterns in the macroeconomic housing market. A logarithmic scale makes the price more manageable.


```{r echo=FALSE}
bronx <- read.csv("~/data/data/rollingsales_bronx.csv")
convertCurrency <- function(currency) {
  currency1 <- sub('$','',as.character(currency),fixed=TRUE)
  currency2 <- as.numeric(gsub('\\,','',as.character(currency1))) 
  currency2
}
bronx <- read.csv("~/data/data/rollingsales_bronx.csv")
bronx$price <- convertCurrency(bronx$SALE.PRICE)
bronx$sq <- as.numeric(bronx$GROSS.SQUARE.FEET)
bronx <- unique(bronx[which(bronx$sq>0 & bronx$price>10000),])
bronx$lsq <-as.numeric(gsub(",","", bronx$LAND.SQUARE.FEET))
bronx$SALE.DATE <- as.Date(gsub("[^]:digit:]]","",bronx$SALE.DATE))
bronx$YEAR.BUILT <- as.numeric(gsub("[^]:digit:]]","",bronx$YEAR.BUILT))
bronx$NEIGHBORHOOD <- as.factor(bronx$NEIGHBORHOOD)
BUILDING.CLASS.CATEGORY <- as.factor(bronx$BUILDING.CLASS.CATEGORY)
```

## Exploration

In the exploration, I looked at the distributions of price and square-foot variables and tentavely tested for a relationship. I decided to use a log-scale for price. I also looked at the log of sale price over time.

```{r}
summary(bronx$price)
summary(bronx$sq)
plot(log(bronx$sq), log(bronx$price))
m1<-lm(log(bronx$price)~log(bronx$sq))
summary(m1)
plot(log(bronx$sq), log(bronx$price))
abline(m1,col="red",lwd=2) 
plot(resid(m1))
bronx$NEIGHBORHOOD <- as.numeric(bronx$NEIGHBORHOOD)
bronx$BUILDING.CLASS.CATEGORY <- as.numeric(bronx$BUILDING.CLASS.CATEGORY)
summary(m1)
plot(log(bronx$sq), log(bronx$price))
abline(m1,col="red",lwd=2) 
plot(resid(m1))
plot(bronx$SALE.DATE, bronx$price)
plot(bronx$SALE.DATE, log(bronx$price))
```

**1.b.** 

It is necessary to clean the data. Currency needs to be converted to a numeric type, as does square feet. It is necessary to filter out 0 price and 0 square feet observations. Neighborhood and building category need to be treated as factors.


    convertCurrency <- function(currency) {
      currency1 <- sub('$','',as.character(currency),fixed=TRUE)
      currency2 <- as.numeric(gsub('\\,','',as.character(currency1))) 
      currency2
    }
    bronx <- unique(read.csv("~/data/data/rollingsales_bronx.csv"))
    bronx <- bronx[which(bronx$sq>0 & bronx$sq>0 & bronx$price>10000),]
    bronx$price <- convertCurrency(bronx$SALE.PRICE)
    bronx$sq <- as.numeric(bronx$GROSS.SQUARE.FEET)
    bronx$lsq <-as.numeric(gsub(",","", bronx$LAND.SQUARE.FEET))
    bronx$SALE.DATE <- as.Date(gsub("[^]:digit:]]","",bronx$SALE.DATE))
    bronx$YEAR.BUILT <- as.numeric(gsub("[^]:digit:]]","",bronx$YEAR.BUILT))
    bronx$NEIGHBORHOOD <- as.factor(bronx$NEIGHBORHOOD)

```{r}
m1<-lm(log(bronx$price)~bronx$sq+factor(bronx$BUILDING.CLASS.CATEGORY)+factor(bronx$NEIGHBORHOOD))
plot(resid(m1))
summary(m1)
```

Our adjusted R-squared is 0.5249, meaning this model explains more than half of the variance in sale price. According to the p-value (p-value < 2.2e-16) this model is highly significant. Many of the building-class coefficients are not significant, but some are highly significant.

**2. a**

## Predicted prices
```{r echo=FALSE}
plot(log(bronx$price))
points(predict.lm(m1, bronx), col="red")
```

The predicted prices look pretty good. The model captured the pattern around the 2000 index very well. A local model would be better at explaining the outliers.

**2. b**  The adjusted R-squared is good, and the model is highly significant (p-value < 2.2e-16). Not all the building classifications are significant.

    Residual standard error: 0.7401 on 3376 degrees of freedom
    Multiple R-squared:  0.5343, Adjusted R-squared:  0.5249 
    F-statistic: 56.97 on 68 and 3376 DF,  p-value: < 2.2e-16

You can see that the model fits the data decently from the adjusted R-squared. The F-statistic is an appropriate test statistic for this model, and it shows that model is significant.

**2. c** There is a lot of data in the dataset we don't want in our model. Those include records that do not represent a sale on the open market. There should be some explanation for the 0-square-feet and no-sale-price properties. It would take research and domain knowledge to go further into it.

**3** The model used was appropriate for learning the global patterns, but a local model would better understand the outliers near the bottom and the top of the price range. A non-linear transform on the square-foot variable might make sense intuitively. Due to the neighborhood and category factors, the model is high-dimensional. More data would make it better and give us some more significant estimators on the some of the coefficients.
